---
title: "Code Portfolio"
author: "Weijian Zhang"
date: "April 13, 2019"
output: html_document
---

```{r setup, include=FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(vcdExtra)
library(vcd)
library(ggplot2)
library(tidyr)
data("Hospital", package ="vcd")
```

## Data Summary

```{r Hospital}
Hospital
```
The Hospital data in vcd gives a 3 x 3 table relating the length of stay (in years) of 132 long-term schizophrenic patients in two London mental hospitals with the frequency of visits by family and friends.


## Data Analysis
To find associate between variables, I perform the Chi-Sq test and to find strength of association, I used assocstats().
```{r chisq test}
chisq.test(Hospital)
assocstats(Hospital)
```
We can conclude that association is significant because the p-values are less than 0.05 for both likelihood ratio and pearson.

## Analysis Visualization
Following is an association plot for these data, with visit frequency as the vertical variable.
```{r assocplot}
assocplot(Hospital)
```
From the above plot, most regular visits last for 2-9 days, and some last for 10-19 days, while less will last for 20+ days.
From frequency aspect, most visits are regular.

##  Extra test to verify conclusion
Because oth variables can be considered ordinal, so CMHtest () may be useful here.
```{r cmtest}
CMHtest(Hospital)
```
Since p-values are less than 0.05, therefore, similar conclusion as association plots.